---
title: "Financial_Analysis"
author: "Amy Cook"
date: "October 21, 2015"
output: word_document
---

#Assess financial implications for a predictive model

We typically get a confusion matrix built from probabilities


```{r loadpackages, echo=FALSE, include=FALSE}

library("plyr")
library("dplyr")
library('magrittr')
library('reshape2')
library('ggplot2')
library('gbm')
library('randomForest')
library('caret')
library('pROC')
library('testthat')
library('purrr')
library(tidyr)
library(broom)

# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
```

Load data 

```{r, echo=FALSE}

test.pred <- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/test_review.csv')
all7d <- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all7d.csv')
test.pred<- merge(test.pred, all7d %>% select(mlsto, balance.mlsto), by = "mlsto")

```


Create vector of profit points


```{r, echo=FALSE}

threshold<- seq(0,1, by=.05)
fin.prof<- rep(0, length(threshold))

for(i in 1:length(threshold)){
        prof<- data.frame("balance" = test.pred$balance.mlsto, 
                  "pred" = test.pred$pred,
                  "ans" = test.pred$b.rpdol)
        #pred = probabilities
        prof$conf <- rep(0, nrow(prof))
        prof$conf <- ifelse(prof$pred >= threshold[i] & prof$ans == 1, 'TP', prof$conf)
        prof$conf <- ifelse(prof$pred >= threshold[i] & prof$ans == 0, 'FP', prof$conf)
        # if model marks it as a positive (greater than threshold), eliminate any balance
        prof$profit<- rep(0, nrow(prof))
        prof$profit<- ifelse(prof$conf %in% "TP", -prof$balance, prof$profit)
        prof$profit<- ifelse(prof$conf %in% "FP", -prof$balance, prof$profit)

        fin.prof[i] = sum(prof$profit)
        
}

prof.plot<- data.frame("threshold" = threshold, "final.profit" = fin.prof)

ggplot(prof.plot, aes(x=threshold, y=final.profit)) + geom_point() + geom_line()

# if say no to all jobs with higher than prob of 0.5, save $500,000

head(prof)


####
# SUMMARY STATS
####

# would dismiss 56 jobs
prof %>% filter(pred>=0.5) %>% nrow

# would dismiss 19 profitable jobs
prof %>% filter(pred >= 0.5 & balance > 0) %>% nrow
#out of 607
prof %>% filter(balance > 0) %>% nrow

# would dismiss 37 losing jobs
prof %>% filter(pred >= 0.5 & balance < 0) %>% nrow
#out of 135
prof %>% filter(balance < 0) %>% nrow

#without algorithm, balance would be $1,475,642
orig.bal<- sum(prof$balance)

# without any losing jobs, balance would be $2,516,246
orig.prof<- prof %>% filter(balance > 0) %>% sum

# without any profitable jobs, balance would be -$1,040,322
orig.loss<- prof %>% filter(balance < 0) %>% sum

#how much real profit would you be dismissing? $82631
prof.dismis<- prof %>% filter(pred >= 0.5 & balance > 0) %>% select(balance) %>% sum
#how much real loss would you be avoiding? $566,481.6
loss.dismis<- prof %>% filter(pred >= 0.5 & balance < 0) %>% select(balance) %>% sum
# total benefit = $483,850
algor.prof<- -(loss.dismis + prof.dismis)

# total balance would be $1,959,492
sum(prof$balance) + algor.prof

# a 32.8% improvement
algor.prof/sum(prof$balance)

```

Write function for making 10 boosted trees and checking the predictions/savings for each!

```{r, echo=FALSE}

all10mice<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all10mice.csv')[,-1]

profit.cv<- function(df.string = "all10mice", method = "boost", seed = 300, k.folds = 3){
        
        results = data.frame('orig.bal' = rep(0, k.folds), 'orig.prof' = rep(0, k.folds), 
                             'orig.loss' = rep(0, k.folds),
                             'prof.dismis' = rep(0, k.folds), 'loss.dismis' = rep(0, k.folds), 
                             'algor.prof' = rep(0, k.folds),
                             'improv' = rep(0, k.folds), 'thresh' = rep(0, k.folds),
                             'auc' = rep(0, k.folds))
        colours <- c("red", "yellow", "green", "blue", "purple")
        
        
        df = get(df.string)
        set.seed(seed)
        
        # create f.rpdol
        df$f.rpdol <- as.factor(df$b.rpdol)
        levels(df$f.rpdol)[levels(df$f.rpdol) == "0"] <-"profit"
        levels(df$f.rpdol)[levels(df$f.rpdol) == "1"] <-"loss"
        
        #set up 10 fold cv using caret function
        folds <- createFolds(df$f.rpdol, k = k.folds, list = T, returnTrain = F)
        #now have folds[[1]] through folds[[10]] list of rows to exclude
        
        
        if(method %in% 'boost') {
                        formula = "b.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + inv.mlsto.log + client.totinv.log + Business + majority.pos + pc.majpos.log"
        } else {
                formula <-"f.rpdol ~ Discipline + pc.pro + b.timespan.cbrt + no.users + inv.mlsto.log + client.totinv.log + Business + majority.pos + pc.majpos.log + JD.Second + Billing.Type"
        }
        
        
        for (j in 1:length(folds)) {
                        
                        #turn predictors and response into 'ordered' variables for roc() to work
                        test<- df[folds[[j]],]
                        levels(test$f.rpdol) <- c('profit', 'loss')
                        test$f.rpdol <- ordered(test$f.rpdol)
                        
                        
                        if(method %in% 'rf'){
                                fit <- randomForest(as.formula(formula), data = df[-folds[[j]],], 
                                            mtry = 5, ntree = 2000)
                                pred <- predict(fit, test, type = "prob")
                                pred <- pred[,2]
                                
                        } else {
                                if(method %in% 'log'){
                                        fit<- glm(as.formula(formula), 
                                                  data = df[-folds[[j]],], family = binomial())
                                        pred <- predict(fit, test,
                                                        type = "response")
                                } else {
                                        fit <- gbm(as.formula(formula), data = df[-folds[[j]],],
                                                   distribution = "bernoulli", n.trees = 10000,
                                                   shrinkage = 0.001, interaction.depth = 4,
                                                   n.minobsinnode = 20)
                                                
                                                pred <-predict(fit, test, 
                                                               n.tree = 10000, type = "response")
                                                }
                        }
                        
                        #pROC
                        pred.p <- roc(test$f.rpdol, pred)
                        # plot.roc(pred.p, col= colours[i], print.thres=T)
                        
                        results$auc[j] <- auc(pred.p)
                        
                        ##
                        #create full test df
                        test$pred<- pred
                        test.pred <- merge(test, all7d %>% select(mlsto, balance.mlsto), by = "mlsto")
                        
                        
                        #create vector of profit points
                        
                        threshold = seq(0,1, by = .05)
                        fin.prof = rep(0, length(threshold))
                        
                        for (i in 1:length(threshold)) {
                                prof <- data.frame("balance" = test.pred$balance.mlsto,
                                "pred" = test.pred$pred,
                                "ans" = test.pred$b.rpdol)
                                
                                #pred = probabilities
                                prof$conf <- rep(0, nrow(prof))
                                prof$conf <- ifelse(prof$pred >= threshold[i] & prof$ans == 1, 'TP', prof$conf)
                                prof$conf <- ifelse(prof$pred >= threshold[i] & prof$ans == 0, 'FP', prof$conf)
                                prof$profit <- rep(0, nrow(prof))
                                prof$profit <- ifelse(prof$conf %in% "TP",-prof$balance, prof$profit)
                                prof$profit <- ifelse(prof$conf %in% "FP",-prof$balance, prof$profit)
                                
                                fin.prof[i] = sum(prof$profit)
                                
                                }
                        
                        prof.plot<- data.frame("threshold" = threshold, "final.profit" = fin.prof)
                        
                        print(
                        ggplot(prof.plot, aes(x=threshold, y=final.profit)) + geom_point() + geom_line()
                        )
                        
                        best.thresh = arrange(prof.plot, -final.profit)[1,1]
                        results$thresh[j] = best.thresh
                        
                        ####
                        # SUMMARY STATS
                        ####
                        
                        #without algorithm, balance would be $1,475,642
                        results$orig.bal[j] = sum(prof$balance)
                        # without any losing jobs, balance would be $2,516,246
                        results$orig.prof[j] = prof %>% filter(balance > 0) %>% sum
                        # without any profitable jobs, balance would be -$1,040,322
                        results$orig.loss[j] = prof %>% filter(balance < 0) %>% sum
                        #how much real profit would you be dismissing? $82631
                        temp<- prof %>% filter(pred >= best.thresh & balance > 0)
                        results$prof.dismis[j] = ifelse(nrow(temp)>0, temp %>% select(balance) %>% sum,
                                                        0)
                        #how much real loss would you be avoiding? $566,481.6
                        results$loss.dismis[j] = prof %>% filter(pred >= best.thresh & balance < 0) %>% 
                                select(balance) %>% sum
                        
                        #print j value
                        cat(j, " ")
                        
        }
        
        # total benefit = $483,850
        results = transform(results, algor.prof = -(loss.dismis + prof.dismis))
        results = transform(results, improv = algor.prof/orig.bal)
        
        return(results)


}

results.1 <- profit.cv(df.string = "all10mice", method = "log", seed = 400, k.folds = 5)
results.2 <- profit.cv(df.string = "all10mice", method = "boost", seed = 400, k.folds =5)
results.3 <- profit.cv(df.string = "all10mice", method = "rf", seed = 300, k.folds = 3)


```


Need a new function for final analysis! comparing two blending methods
Can i get rid of boosted tree model? that would speed things up a lot

Need to try simple average and log reg without boosted trees. Reduces spread but doesnt improve mean. i'd like to keep the wide spread because you reach greater heights (and lower lows)

```{r}
if(.Platform$OS.type == 'windows'){
        all10mice = read.csv('C:/Users/n9232371/OneDrive/shared files/Bligh Tanner/masters/data/all10mice_May16.csv')[,-1]
        source("C:/Users/n9232371/Documents/github/consult/functions.R")
} else{
        all10mice = read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all10mice_May16.csv')[,-1]
        source("/Users/yam/Documents/github/consult/functions.R")

}

# #clean up some JD.SEcond categories!
# summary(all10mice$JD.Second)
# # all Subdivision sohould be under civBldg_subdiv_sewer
# all10mice$JD.Second <- as.character(all10mice$JD.Second)
# all10mice$JD.Second <- ifelse(all10mice$JD.Second == "Subdivision", "civBldg_subdiv_sewer",
#                               all10mice$JD.Second)
# all10mice$JD.Second <- as.factor(all10mice$JD.Second)
# # mlsto 20090.238 should be flood harvest
# all10mice[all10mice$mlsto == "2009.238",]$JD.Second <- 'flood_h2o_harvest'
# write.csv(all10mice, '~/OneDrive/shared files/Bligh Tanner/masters/data/all10mice_May16.csv')


profit <- function(df = all10mice, method = c("average","simp.log","simp.boost","orig.log","orig.rf","orig.boost"), 
                   seed = 400, 
                   k.folds = 5, # number of folds in initial model.blend generation
                   max.iter = 5, #max number of iterations from initial model.blend generation
                   k.folds.2 = 5, # number of folds for second step blended model development
                   tot.tests = 500 # total number of results to create
                   ){
        
        length.k.folds = ceiling(tot.tests/k.folds.2)*k.folds.2
        
        set.seed(seed)
        
        # create f.rpdol
        df$f.rpdol <- as.factor(df$b.rpdol)
        levels(df$f.rpdol)[levels(df$f.rpdol) == "0"] <-"profit"
        levels(df$f.rpdol)[levels(df$f.rpdol) == "1"] <-"loss"
        
        # testthat check
        expect_equal(levels(df$f.rpdol), c("profit", "loss"))
        
        #loop through however many versions of df.blend are required
        compile.results = NULL
        
        #store results of three original models in list for each iteration
        joined <- assemble(methods = c("log", "rf", "boost"), df.assem = df, x.folds.assem = 5,
                    iterations = ceiling(tot.tests/k.folds.2))
        
        # saveRDS(joined, "/Users/yam/Documents/github/consult/finalwriteup/report_data/joined2.rds")
        # run blended models on each of the 100 outputs from original models stored
        # in joined list. Store as new list 'compile.results' for each blending method
        for(m in 1:ceiling(tot.tests/max.iter)){
        
        print(
                paste("m =", m, "of", ceiling(tot.tests/max.iter),"\n", sep = ' ') %>% cat()
        )        
                
        # create complete set using mth list item in joined
        df.blend = full_join(df %>% mutate(rown = 1:nrow(df)), joined[[m]], 
                                by = c('rown'='rownames'))
        df.blend <- df.blend %>% select(-rown)
        
        # df.blend = assemble(methods = c("log","rf","boost"), df.assem = df, x.folds.assem = 5)
        
        
        df.blend = mutate(df.blend, norm.log = sign(log) * abs(log)^(1/3))
        df.blend = mutate(df.blend, norm.rf = rf^0.5)
        df.blend = mutate(df.blend, norm.boost = log(boost))
        
        # saveRDS(df.blend, "C:/Users/n9232371/Documents/github/consult/dfblend.rds")
        # df.blend <- readRDS("/Users/yam/Documents/github/consult/dfblend.rds")
        
        saveit <- iter.auc(mult = max.iter/k.folds.2, x.folds.iter = k.folds.2,
                           methods = c(
                                  "simp.log","average",
                                  "orig.log","orig.rf","orig.boost",
                                  "comp.rf","comp.boost","FWLS",
                                   "simp.boost"
                                  ),
                           df.iter = df.blend)

        dontknow = map(saveit, function(x) spread(x, iteration, pred))
        
        if(m == 1){
                compile.results = dontknow
        } else {
                compile.results = map2(compile.results, dontknow, function(x,y) merge(x, y, by = 'index'))
        }
        
        }
        
        # saveRDS(compile.results, "/Users/yam/Documents/github/consult/finalwriteup/report_data/finance_results.rds")
        # compile.results <- readRDS("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/finance_results.rds")
        
        
        ####
        # now have a suite of predictions - merge with original all10.blend$balance.mlsto
        df.blend$index = rownames(df.blend) %>% as.numeric
        # load balances
        if(.Platform$OS.type == 'windows'){
        all6 <-read.csv('C:/Users/n9232371/OneDrive/shared files/Bligh Tanner/masters/data/all6.csv')
        } else{
        all6 <-read.csv('/Users/yam/OneDrive/shared files/Bligh Tanner/masters/data/all6.csv')
        }
        #iOS
        all6 <- all6 %>% select(mlsto, balance.mlsto) %>% distinct()
        all6$mlsto <- as.character(all6$mlsto)
        df.blend = left_join(df.blend, all6)
        compile.results <- map(compile.results, function(x) 
                left_join(x, df.blend %>% select(index, balance.mlsto, b.rpdol))) 

                        
        #create vector of profit points
        threshold = seq(0, 1, by = .05)

        # set up list of blank dataframes for profit calcs for each iteration: prof.results
        colnames = names(compile.results[[1]])[grepl('iter', names(compile.results[[1]]))]
        new.df = matrix(NA, nrow = length(threshold), ncol = ceiling(tot.tests/max.iter) + 1) %>% as.data.frame
        colnames(new.df) <- c('threshold', colnames)
        new.df$threshold = threshold 
        prof.results = list(new.df,new.df,new.df,new.df,new.df,new.df,new.df,new.df,new.df)
        names(prof.results) = names(compile.results)
        
         
        # try eliminating outliers of balance over $200,000
        # comp.results.2 <- lapply(compile.results, function(x) x = x %>% filter(abs(balance.mlsto)<200000))
        
        prof.results <- map2(compile.results, prof.results, function(x,y) {
                
                for(n in 2:ncol(y)){
                        
                        for (p in 1:length(threshold)) {
                        
                        # sum the balance.mlsto of all jobs that fall below the cut off threshold
                        # this equals the total profits (or losses!) across all jobs that were completed
                        # given that you reject any job over the threshold
                                
                        prof <- x[x[,n] < y$threshold[p], 'balance.mlsto']
                                
                        y[p, n] = sum(prof)
                                
                }
                        
                }
                
                return(y)
                
        })
        
        prof.results <- map2(prof.results, names(prof.results), function(x,y) {
                x$method = y
                return(x)})
        
        assign("prof.results", prof.results, envir = .GlobalEnv)
                        
}

#function below assigns results to 'prof.results' object in global env
profit(df = all10mice, method = c("average","simp.log",'simp.boost','FWLS','comp.boost','comp.rf', "orig.log","orig.rf","orig.boost"),
                   seed = 300,
                   k.folds = 5, # number of folds in initial model.blend generation
                   max.iter = 5, #max number of iterations from initial model.blend generation
                   k.folds.2 = 5, # number of folds for second step blended model development
                   tot.tests = 500 # total number of results to create
                   )

# gather results for plotting
plot.results = rbind_all(prof.results)
plot.results = gather(plot.results, key = iteration, value = profit, -method, - threshold)

# censor actual profit amount as fraction of historical profit level
orig.prof = plot.results %>% filter(threshold == 1) %>% .$profit %>% .[1]
plot.results = mutate(plot.results, profit.ratio = (profit/orig.prof*100) %>% round(1))

saveRDS(plot.results,"C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/plot_finance_100_4.rds" )
saveRDS(plot.results,"/Users/yam/Documents/github/consult/finalwriteup/report_data/plot_finance_100_4.rds" )

bind <- bind %>% filter(!method %in% 'FWLS')
bind <- rbind(bind, plot.results)

ggplot(plot.results, aes(x= threshold, y= profit.ratio, colour = method)) + 
        geom_point(alpha = 0.1, shape = 16) + 
        stat_smooth() +
        scale_y_continuous(breaks = seq(0,110, by = 10)) + 
        scale_x_continuous(breaks = seq(0,1, by = .1)) + 
        geom_hline(yintercept = 100, colour = "darkgray", linetype = 2)


# why is it so shit? calc average prob for each iteration
mean.prob = apply(compile.results[['average']], 1, function(x) mean(x[2:11])) 
analysis = data.frame(index = 1:length(mean.prob), mean.prob = mean.prob, 
                      bal = compile.results[['simp.log']]$balance.mlsto)
analysis %>% arrange(mean.prob) %>% head(10)
ggplot(analysis, aes(x = mean.prob, y = bal)) + geom_point()
        
```


Now need to check how mnay experiments to do.. power calc. 

```{r}

#power calc - where should I do it from? 

```

Check for significant difference between samples

```{r t_test}
plot.results

index = c(2:8)
pvals = data.frame("method" = levels(plot.results$method %>% as.factor)[c(2:8)], 
                   "threshold" = 
                   'p_value' = NA)
for(i in seq_along(index)){
        pvals$p_value[i] = t.test(auc.150[[1]], 
                                  auc.150[[index[i]]])$p.value %>%
                format(digits = 2) %>% as.numeric
}

```





Check what happens when you increase the amount you invoiced by a certain percentage

```{r}

# what if you find for jobs over a certain threshold, increase fee by ?% still do all jobs
compile.results <- readRDS("C:/Users/n9232371/Documents/github/consult/finalwriteup/report_data/finance_results.rds")
####
# now have a suite of predictions - merge with original all10.blend$balance.mlsto
df.blend$index = rownames(df.blend) %>% as.numeric
# load balances
all6 <-read.csv('C:/Users/n9232371/OneDrive/shared files/Bligh Tanner/masters/data/all6.csv')
all6 <- all6 %>% select(mlsto, balance.mlsto, inv.mlsto, cost.mlsto) %>% distinct()
all6$mlsto <- as.character(all6$mlsto)
df.blend = left_join(df.blend, all6)
compile.results <- map(compile.results, function(x) {
        left_join(x, df.blend %>% select(index, balance.mlsto, b.rpdol, inv.mlsto, cost.mlsto))}) 
                        
#create vector of profit points
threshold = seq(0, 1, by = .05)

# set up list of blank dataframes for profit calcs for each iteration: prof.results
colnames = names(compile.results[[1]])[grepl('iter', names(compile.results[[1]]))]
new.df = matrix(NA, nrow = length(threshold), ncol = ceiling(tot.tests/max.iter) + 1) %>% as.data.frame
colnames(new.df) <- c('threshold', colnames)
new.df$threshold = threshold 
prof.results = list(new.df, new.df, new.df, new.df, new.df)
names(prof.results) = names(compile.results)

        
prof.results <- map2(compile.results, prof.results, function(x,y) {
                
        for(n in 2:ncol(y)){
                        
                for (p in 1:length(threshold)) {
                        
                # sum the balance.mlsto of all jobs - you do them all
                # then sum a 10% increase in what you charge for all jobs above a certain threshold
                                
                        prof <- x[, 'balance.mlsto']
                        increase <- x[x[,n] > y$threshold[p], 'inv.mlsto']*0.1
                                
                        y[p, n] = sum(prof) + sum(increase)
                                
                }
                        
        }
                return(y)
})
        
prof.results <- map2(prof.results, names(prof.results), function(x,y) {
        x$method = y
        return(x)})
        
assign(prof.results, "prof.results", envir = .GlobalEnv)


# gather results for plotting
plot.results = rbind_all(prof.results)
plot.results = gather(plot.results, key = iteration, value = profit, -method, - threshold)

ggplot(plot.results, aes(x= threshold, y= profit, colour = method)) + 
        geom_point(alpha = 0.1, shape = 16) + 
        stat_smooth()

# doesn't reveal much - better off doing a simple check once decided threshold.

# I like simp.log best. 

simp.l <- compile.results[['simp.log']]
simp.l$mean.iter <- rowMeans(simp.l[,2:101])

saveRDS(simp.l, "/Users/yam/Documents/github/consult/finalwriteup/report_data/finance_average_model.rds")

# How many jobs are above threshold of 0.65?
simp.l %>% filter(mean.iter>0.65) %>% nrow
# 74 - not many

#how much was lost above 0.65?
initial = simp.l %>% filter(mean.iter>0.65) %>% .$balance.mlsto %>% sum
# $600000
# what if increased invoice amount by 0.1?
costs = simp.l %>% filter(mean.iter>0.65) %>% .$cost.mlsto %>% sum
simp.l %>% filter(mean.iter>0.65) %>% mutate(frac = inv.mlsto*0.27) %>% .$frac %>% sum

# would have to increase all the jobs which rated above 0.65 by 27% to break even.
# would have to increase all the jobs which rated above 0.65 by 46% to generate a 15% profit.
# this is probably impractical - hard to compete or argue for it. could also go for Hourly rate billing

profit.15 = costs*.15


```








